---
title: "Case study analysis"
author: "Eddie Chapman"
date: "2/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r libraries}
library(RColorBrewer)
library(tidyverse)
library(treemap)
```
```{r import_data}
survey <- read_rds('data/survey_data.rds')
```

# Data Cleaning

Filter columns down to respondant ID, key factors, and IRB determination for each case study.

Transform the data into long form as: ID, case study #, irb determination, key factor (one factor chosen by one respondent for this case study).

I separated the original survey data into a `determination` dataframe and a `key_factors` dataframe. I was having trouble with `pivot_longer` creating a duplicate `case_study` column using the various column labels. 

```{r data_cleaning}
determination <- survey %>%
  select(id, ends_with("irb_would_consider_proposal")) %>%
  pivot_longer(ends_with("irb_would_consider_proposal"), 
               names_to = "case_study",
               names_pattern = "c(\\d+)_.+",
               values_to = "irb_determination",
               values_drop_na = TRUE) %>%
  mutate_if(is.character, as.factor)

key_factors <- survey %>%
  select(id, contains("key_factor"), -contains("other_text")) %>%
  pivot_longer(contains("key_factor"),
               names_to = c("case_study", "key_factor"),
               names_pattern = "c(\\d+)_key_factor_(.*)",
               values_to = "response",
               values_drop_na = TRUE) %>%
  filter(response == TRUE) %>%
  select(-response) %>%
  mutate_if(is.character, as.factor)
```

Join them into a single dataframe and discard the unused dataframes.

```{r}
case_studies <- determination %>%
  left_join(key_factors, by = c("id", "case_study"))

rm(key_factors, determination)

case_studies
```

The `NA` value in row 1 needs to be addressed. `NA`s indicate that a respondent filled out the `irb_determination` question without choosing any key factors. An option of "other" should be chosen in this situation. 

We'll remove the `NA`s in the `key_factor` column.

```{r data_cleaning_NAs}
# NAs in key_factor means that respondant answered irb_determination
# question but did not offer any key factors. Dropping these.
case_studies <- case_studies %>%
  filter(!is.na(key_factor))
```

## Analysis

Top 5 factors and their counts per determination, per case study

```{r}
tree <- case_studies %>% 
  select(-id) %>%
  mutate(
    irb_determination = fct_recode(
      irb_determination,
      "Exempt" = "exempt",
      "Expedited" = "expedited",
      "Not human subjects research" = "not human subjects research",
      "Full review" = "full review"),
    key_factor = fct_recode(
      key_factor,
       "Public vs. private site" = "public_or_private_site",
       "Public vs. private data" = "public_or_private_data",
       "Level of analysis" = "level_of_analysis",
       "Identifiable data" = "pid",
       "Terms of service" = "tos",
       "Data reuse" = "data_reuse",
       "Combines datasets" = "combine_dataset",
       "Method of obtaining data" = "method_obtain_data",
       "Type of data" = "type_of_data",
       "Obtains informed consent" = "informed_consent",
       "Purpose of the research" = "research_purpose",
       "Impact beyond participants" = "research_impact",
       "Other" = "other")) %>%
  add_count(case_study, name = "n_case") %>%
  add_count(case_study, irb_determination, name = "n_determination") %>%
  add_count(case_study, irb_determination, key_factor, name = "n_factor") %>%
  select(
    case_study, 
    n_case, 
    irb_determination, 
    n_determination, 
    key_factor, 
    n_factor) %>%
  arrange(
    case_study, 
    desc(n_determination), 
    irb_determination, 
    desc(n_factor), 
    key_factor) %>%
  group_by(case_study, irb_determination) %>%
  top_frac(.5, wt = n_factor) %>%
  count(case_study, irb_determination, key_factor) %>%
  arrange(
    case_study, 
    irb_determination, 
    desc(n), 
    key_factor)

tree
  
```

Top 5 factors and their counts per determination

```{r}
case_studies %>%
  count(irb_determination, key_factor)
```

```{r}
case_studies %>%
  select(case_study, key_factor, irb_determination, -id) %>%
  add_count(case_study, key_factor, name = "per_case") %>%
  add_count(case_study, irb_determination, key_factor, 
            name = "per_determination") %>%
  mutate(percent = per_determination / per_case) %>%
  unique() %>%
  arrange(case_study, key_factor) 
```

## Treeplots

This is a way to visualize the breakdown between IRB determinations for each case study and the factors that influenced them. 


Per case study, the percentage of each key factor that was included in each determination

Not looking at how many people chose a given determination - just what caused them to do so.

Need:

- # of key factors chosen per determination per case study
- # of times a given factor was chosen per determination per case study
- # of times a given factor was chosen per case study

```{r data_for_treeplots}
# tree <- case_studies %>%
#   select(case_study, irb_determination, key_factor, -id) %>%
#   add_count(case_study, key_factor, name = "per_case") %>%
#   add_count(case_study, irb_determination, key_factor, 
#             name = "per_determination") %>%
#   mutate(percent = per_determination / per_case,
#          key_factor = fct_recode(
#            key_factor,
#            "Public vs. private site" = "public_or_private_site",
#            "Public vs. private data" = "public_or_private_data",
#            "Level of analysis" = "level_of_analysis",
#            "Identifiable data" = "pid",
#            "Terms of service" = "tos",
#            "Data reuse" = "data_reuse",
#            "Combines datasets" = "combine_dataset",
#            "Method of obtaining data" = "method_obtain_data",
#            "Type of data" = "type_of_data",
#            "Obtains informed consent" = "informed_consent",
#            "Purpose of the research" = "research_purpose",
#            "Impact beyond participants" = "research_impact",
#            "Other" = "other")) %>%
#   unique() %>%
#   arrange(case_study, irb_determination, desc(per_determination)) %>%
#   group_by(case_study, irb_determination) %>%
#   top_frac(0.5, wt = per_determination)
# 
# tree
```

Saving the titles as variables for use in plotting.

```{r case_study_titles}
title1 <- "1. Scrapping public newspaper comments to predict elections"

title2 <- "2. Scrapping public Twitter feeds to predict risky drug-use behaviors"

title3 <- "3. Analyzing dating site data to infer sexual behavior"

title4 <- "4. Analyzing newspaper comments to understand political views"

title5 <- "5. Collect geolocation data from mobile provider to understand group mobility patterns"

title6 <- "6. Combine mental health data with social media activity"

title7 <- "7. Analyzing preexisting Twitter dataset to study political event"

title8 <- "8. Scraping health forum and combining with Twitter data to predict mental health"

title9 <- "9. Scraping profile photos to predict sexual behavior"

title10 <- "10. Analyze Apple HealthKit data to assess impact of exercise on health"

title11 <- "11. Scrape public Facebook posts to study group-level dynamics"
```

```{r add_color_to_determination}
palette = tibble(
  color = c(
    brewer.pal(6, "Pastel1")[3],
    brewer.pal(6, "Pastel1")[6],
    brewer.pal(6, "Pastel1")[2],
    brewer.pal(6, "Pastel1")[1]),
  irb_determination = c(
    "Exempt", 
    "Expedited", 
    "Not human subjects research", 
    "Full review"))

tree <- tree %>%
  left_join(palette, by = "irb_determination")

tree
```


```{r plot_treemap_func}
#
# Plot and save a treemap for a single case study in the dataset
#
plot_treemap <- function(filename, data, case_number, title) {
  
  ppi <- 300
  
  png(filename, 
      width = 7 * ppi, 
      height = 4 * ppi, 
      res = ppi)
  
  tree <- treemap(data %>% filter(case_study == case_number),
                  index = c("irb_determination", "key_factor"),
                  vSize = "n",
                  vColor = "color",
                  type = "color",
                  title = title,
                  title.legend = "Key factor",
                  bg.labels = 180,
                  fontsize.labels = c(12, 8),
                  fontsize.title = 13,
                  fontsize.legend = 10,
                  position.legend = "none",
                  fontface.labels = c("bold.italic", "plain"),
                  align.labels = list(
                    c("left", "top"),
                    c("center", "center")
                  ),
                  inflate.labels = FALSE,
                  palette = "Pastel1",
                  overlap.labels = 0.25,
                  border.col = c("black", "black"),
                  border.lwds = c(2.5, 0.5),
                  lowerbound.cex.labels = 0.75,
                  range = c(0, 1),
                  aspRatio = 1.5)
  
  dev.off()
}
```


```{r plot_trees_and_save}
plot_treemap(filename = "plots/case-01_news-comments-election.png",
             data = tree, case_number = 1, title = title1)

plot_treemap(filename = "plots/case-02_twitter-drug-use.png",
             data = tree, case_number = 2, title = title2)

plot_treemap(filename = "plots/case-03_dating-sites.png", 
             data = tree, case_number = 3, title = title3)

plot_treemap(filename = "plots/case-04_news-comments-political-views.png",
             data = tree, case_number = 4, title = title4)

plot_treemap(filename = "plots/case-05_group-mobility-patterns.png",
             data = tree, case_number = 5, title = title5)

plot_treemap(filename = "plots/case-06_mental-health-social-media.png",
             data = tree, case_number = 6, title = title6)

plot_treemap(filename = "plots/case-07_preexisting-twitter-political-event.png",
             data = tree,case_number = 7, title = title7)

plot_treemap(filename = "plots/case-08_health-forum-plus-twitter-mental-health.png",
             data = tree, case_number = 8, title = title8)

plot_treemap(filename = "plots/case-09_profile-pics-sexual-behavior.png",
             data = tree, case_number = 9, title = title9)

plot_treemap(filename = "plots/case-10_apple-health-kit.png",
             data = tree, case_number = 10, title = title10)

plot_treemap(filename = "plots/case-11_public-facebook-group-dynamics.png",
             data = tree, case_number = 11, title = title11)
```
