---
title: "Survey Data Visualization"
output:
  html_notebook: default
  pdf_document: default
---

```{r}
library(ggthemes)
library(knitr)
library(RColorBrewer)
library(tidyverse)
```


# Functions

```{r}
bar_plot <- function(df, field, title, subtitle) {
  
  field <- enquo(field)
  
  if(missing(title)) { 
    title = field 
  }
  
  if(missing(subtitle)) {
    subtitle = ""
  }
  
  df %>% 
    drop_na(!! field) %>%
    ggplot(aes(!! field)) +
    geom_bar() +
    guides(fill = FALSE, color = FALSE) +
    labs(x = NULL, 
         y = "Responses",
         title = title, 
         subtitle = subtitle)
}


text_responses <- function(df, field) {
  
  field <- enquo(field)
  
  df %>%
    select(!! field) %>%
    drop_na(!! field) %>%
    kable()
}

```

```{r}
survey <- read_rds('data/survey_data.rds')
head(survey)
```
# Survey

## Institution

### Organization type

```{r}
bar_plot(survey, org_type, title = "Organization Type", subtitle = )
```

"Other" responses:

```{r echo = FALSE}
text_responses(survey, org_type_text)
```

### IRB ethics primary concern

```{r}
bar_plot(survey, 
         irb_ethics_primary_concern, 
         title = "IRB Ethics Concerns", 
         subtitle = "My IRB/ethics concerns are primarily focused on...")
```

"Other" responses:
```{r echo = FALSE}
text_responses(survey, irb_ethics_primary_concern_text)
```

### Academic institution type

```{r}
bar_plot(survey, academic_institution_type, "Academic Institution Type")
```

## Personal Experience

### Roles

```{r}

```

### Academic research type

```{r}
bar_plot(survey, 
         academic_research_type, 
         title = "Academic Research Type", 
         subtitle = "The type of research I primarily review/oversee is primarily...")
```

### Annual pervasive protocols


```{r}
bar_plot(survey, 
         annual_pervasive_protocols, 
         title = "Number of pervasive protocols reviewed", 
         subtitle = "How many protocols involving pervasive data does your IRB review annually?")
```


### Percentages 

```{r}

```


### Reviewed pervasive data types

```{r}

```

"Other" text:

```{r}
text_responses(survey, reviewed_other_text)
```

### Most pervasive data disciplines

```{r}
text_responses(survey, most_pervasive_data_in_other_text)
```


## IRB Policy

### Confidence likert questions

Set up the color palette.

```{r}
pal = brewer.pal(5, "RdBu")
pal[3] <- "#DFDFDF"  # A little brighter

likert <- c("strongly disagree", 
            "somewhat disagree", 
            "neither agree nor disagree",
            "somewhat agree", 
            "strongly agree")
                 

palette <- tibble(response = factor(likert, levels = likert, ordered = TRUE),
                  color = factor(pal, ordered = TRUE))
```

Divide the responses into `high` and `low` dataframes. Half of the "neither agree nor disagree" responses will be in each. This allows the middle point to be centered for comparison between questions.

```{r}
confidence <- survey %>%
  select("irb_members_ethical" = "confidence_irb_members_pervasive_data_ethical",
         "irb_members_technical" = "confidence_irb_members_pervasive_data_technical",
         "researchers_ethical" = "confidence_researchers_pervasive_data_ethical") %>%
  pivot_longer(cols = irb_members_ethical:researchers_ethical, 
               names_to = "consideration", 
               names_ptypes = list("consideration", factor()),
               values_to = "response",
               values_ptypes = list("response", factor())) %>%
  drop_na() %>%
  group_by(consideration, response) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  left_join(palette, by = c('response'))

confidence_low <- confidence %>%
  filter(response %in% levels(response)[1:3]) %>%
  mutate(freq = ifelse(str_detect(response, levels(response)[3]), freq / 2, freq),
         freq = as.integer(round(freq * 100), 2),
         response = factor(response),
         color = factor(color, levels = c("#CA0020", "#F4A582", "#DFDFDF")))

confidence_high <- confidence %>%
  filter(response %in% levels(response)[3:5]) %>%
  mutate(freq = ifelse(str_detect(response, levels(response)[3]), freq / 2, freq),
         freq = as.integer(round(freq * 100), 2))
```

#### All three likert questions

Plot likert data for all three question responses.

```{r}
title <- "Prepared for pervasive data protocols?"
subtitle <- "I believe that (IRB members, researchers) at my institution are well-versed\nin the (technical, ethical) aspects of the collection/use of pervasive data."

p <- ggplot() +
  
  geom_bar(data = confidence_high,
           aes(x = factor(consideration), y = freq, fill = factor(color)),
           position = "stack", 
           stat = "identity", 
           width = 0.5) +
  
  geom_bar(data = confidence_low,
           aes(x = factor(consideration), y = -freq, fill = factor(color)),
           position = "stack", 
           stat = "identity", 
           width = 0.5) +
  
  geom_hline(yintercept = 0, color = c("white")) +
  
  scale_x_discrete(breaks = c("researchers_ethical", "irb_members_technical", "irb_members_ethical"),
                   labels = c("Researchers,\nethical", "IRB members,\ntechnical", "IRB members,\nethical")) +
  
  scale_fill_identity(name = "Percent respondents (n = 76)", 
                      labels = rev(c("Strongly disagree", 
                                     "Somewhat disagree", 
                                     "Neither agree nor disagree", 
                                     "Somewhat agree", 
                                     "Strongly agree")), 
                      breaks = rev(palette$color), 
                      guide = "legend") +
  
  theme_fivethirtyeight() +
  
  coord_flip() +
  
  labs(title = title, 
       subtitle = subtitle, 
       y = "", 
       x = "Percent respondents (n = 76)") +
  
  theme(plot.title = element_text(size = 14, hjust = 0),
        plot.subtitle = element_text(size = 12, face = "italic", hjust = 0),
        axis.text.y = element_text(size = 11, face = 'bold', hjust = .5),
        legend.position = "right",
        legend.direction = "vertical",
        legend.key.size = unit(.4, "cm"),
        legend.text = element_text(size = 11)) +
  
  scale_y_continuous(breaks = seq(-100 , 100, 25), 
                     limits = c(-75, 75))
```

Save the plot to device.

```{r}
ppi <- 300
png("plots/likert_plot_with_researchers.png", width = 7 * ppi, height = 4 * ppi, res = ppi)
p
dev.off()
```

#### Without "researchers" question

Again but removing the "researcher" question. Removing from dataframe:

```{r}
title <- "Prepared for pervasive data protocols?"
subtitle <- "I believe that IRB members at my institution are well-versed in the\n(technical, ethical) aspects of the collection/use of pervasive data."

p2 <- ggplot() +
  
  geom_bar(data = filter(confidence_high, consideration != "researchers_ethical"),
           aes(x = factor(consideration), y = freq, fill = factor(color)),
           position = "stack", 
           stat = "identity", 
           width = 0.5) +
  
  geom_bar(data = filter(confidence_low, consideration != "researchers_ethical"),
           aes(x = factor(consideration), y = -freq, fill = factor(color)),
           position = "stack", 
           stat = "identity", 
           width = 0.5) +
  
  geom_hline(yintercept = 0, color = c("white")) +
  
  scale_x_discrete(breaks = c("irb_members_technical", "irb_members_ethical"),
                   labels = c("Technical", "Ethical")) +
  
  scale_fill_identity(name = "Percent respondents (n = 76)", 
                      labels = rev(c("Strongly disagree", 
                                     "Somewhat disagree", 
                                     "Neither agree nor disagree", 
                                     "Somewhat agree", 
                                     "Strongly agree")), 
                      breaks = rev(palette$color), 
                      guide = "legend") +
  
  theme_fivethirtyeight() +
  
  coord_flip() +
  
  labs(title = title, 
       subtitle = subtitle, 
       y = "", 
       x = "Percent respondents (n = 76)") +
  
  theme(plot.title = element_text(size = 14, hjust = 0),
        plot.subtitle = element_text(size = 12, face = "italic", hjust = 0),
        axis.text.y = element_text(size = 11, face = 'bold', hjust = .5),
        legend.position = "right",
        legend.direction = "vertical",
        legend.key.size = unit(.4, "cm"),
        legend.text = element_text(size = 11)) +
  
  scale_y_continuous(breaks = seq(-100 , 100, 25), 
                     limits = c(-75, 75))

ppi <- 300
png("plots/likert_plot_no_researchers.png", width = 7 * ppi, height = 4 * ppi, res = ppi)
p2
dev.off()
```


### IRB has policy for pervasive data

```{r}
q <- "Does your IRB have a specific checklist, review tool, policy or\nguidelines for reviewing protocols that rely on pervasive data?"

bar_plot(survey, 
         irb_has_policy_for_pervasive_data,
         title = "IRB Policy",
         subtitle = q)
```


```{r}
bar_plot(survey, 
         irb_policy_for_pervasive_data_usefulness,
         title = "IRB policy usefulness",
         subtitle = "How would you evaluate the usefulness of these policies or documents?")
```

### IRB Guidelines


```{r}
text_responses(survey, irb_guidelines_other_text)
```


### Institution trains researchers/IRB members

```{r}
bar_plot(survey, 
         institution_trains_researchers_pervasive_data_ethics,
         title = "Training for Researchers",
         subtitle = "Does your institution provide specific training sessions for RESEARCHERS\nthat addresses the ethics of the collection/use of pervasive data?")
```

```{r}
bar_plot(survey, 
         institution_trains_irb_members_pervasive_data_ethics,
         title = "Training for IRB Members",
         subtitle = "Does your institution provide specific training sessions for IRB MEMBERS\nthat addresses the ethics of the collection/use of pervasive data?")
```

### IRB implemented changes to common rule

```{r}
bar_plot(survey, 
         irb_implemented_changes_to_common_rule,
         title = "Common Rule Implemented",
         subtitle = "Has your IRB already implemented changes to the Common Rule?")
```

## Case Studies

